A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Additional Methods

### 2020-12-06 writeUp/supplementary_materials_methods.tex
very nearly at a draft

\subsubsection{Impact of insecticide interventions on lifespan}\label{sec:detinova_insecticide}
To determine the impact of insecticidal interventions on lifespan, we modified the model described in \S\ref{sec:detinova_datamodel} to estimate effects at the species level. To do so, we allowed a logit-linear impact of insecticide. Further, the species-level effects were modelled hierarchically using normal priors, allowing the impact of resistance to vary at the observation level within each study. The model to estimate this is shown below (where the variable metering the impact of insecticide at the observational level is `delta').

@@

To investigate the effect of measures of temperature, we extracted daily data from the ERA-interim re-analysis by ECMWF \citep{dee2011era} through their website using the Python API. This data included:

For this analysis, we included only those studies where the start month listed in the database coincided with the end date - meaning the study took place within a given month - resulting in $n=742$ data points. The temperature data was extracted for the GPS position of each study ($\pm 0.1 \text{ degrees}$) for each day of the month over which mosquito sampling took place, with a grid size of 0.05 degrees. The Python script to extract this information is given below.

\begin{itemize}
	\item Average two meter temperature (average of mean daily data for each grid point),
	\item Minimum two meter temperature (average of minimum daily data for each grid point),
	\item Maximum two meter temperature (average of maximum daily data for each grid point),
	\item Standard deviation in two meter temperature (average of standard deviation in temperature for each grid point),
	\item Daily range in two meter temperature (average of maximum minus minimum daily temperature for each grid point).
\end{itemize}

@@

>>>>

@@

@@

\subsection{Association of insecticide interventions on lifespan}\label{sec:detinova_insecticide}
To determine the association of insecticidal interventions with lifespan, we modified the model described in \S\ref{sec:detinova_datamodel} to estimate effects at the species level. To do so, we allowed a logit-linear impact of insecticide. Further, the species-level effects were modelled hierarchically using normal priors, allowing the impact of resistance to vary at the observation level within each study. The model to estimate this is shown below (where the variable metering the impact of insecticide at the observational level is `delta'). This model was estimated using 1000 iterations per chain (500 of which were warm-up) across 12 chains.

\subsection{Association of weather covariates with temperature}
To investigate the effect of measures of temperature, we extracted daily data from the ERA-interim re-analysis by ECMWF \citep{dee2011era} through their website using the Python API. This data included:

For this analysis, we included only those studies where the start month listed in the database coincided with the end date, meaning the study took place within a given month. We also omitted those data points where insecticide use was recorded. To ensure that we had sufficient data points for each fold used for model comparison, we considered only those species with over 100 observations. This resulted in $n=366$ observations across 2 species complexes: \textit{A. gambiae s.l.} ($n=257$) and \textit{A. funestus s.l.} ($n=109$).

The temperature data was extracted for the GPS position of each study ($\pm 0.1 \text{ degrees}$) for each day of the month over which mosquito sampling took place, with a grid size of 0.05 degrees. The Python script to extract this information is given below.

\begin{itemize}
	\item Average two meter temperature (average of mean daily data for each grid point),
	\item Daily range in two meter temperature (average of maximum minus minimum daily temperature for each grid point). This likely reflects the average day-night temperature range.
\end{itemize}

Each of these two models were estimated in Stan using 1000 iterations per chain (500 of each were warm-up) across 4 chains and thinned samples by a factor of 20. To determine whether each covariate was predictive for lifespan, we undertook K-fold cross-validation (using 4 folds with a given fold holding roughly equal numbers of each species).
